## Introduction

This assignment uses data from
the <a href="http://archive.ics.uci.edu/ml/">UC Irvine Machine
Learning Repository</a>, a popular repository for machine learning
datasets. In particular, we will be using the "Individual household
electric power consumption Data Set" which I have made available on
the course web site:


* <b>Dataset</b>: <a href="https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2Fhousehold_power_consumption.zip">Electric power consumption</a> [20Mb]

* <b>Description</b>: Measurements of electric power consumption in
one household with a one-minute sampling rate over a period of almost
4 years. Different electrical quantities and some sub-metering values
are available.


The following descriptions of the 9 variables in the dataset are taken
from
the <a href="https://archive.ics.uci.edu/ml/datasets/Individual+household+electric+power+consumption">UCI
web site</a>:

<ol>
<li><b>Date</b>: Date in format dd/mm/yyyy </li>
<li><b>Time</b>: time in format hh:mm:ss </li>
<li><b>Global_active_power</b>: household global minute-averaged active power (in kilowatt) </li>
<li><b>Global_reactive_power</b>: household global minute-averaged reactive power (in kilowatt) </li>
<li><b>Voltage</b>: minute-averaged voltage (in volt) </li>
<li><b>Global_intensity</b>: household global minute-averaged current intensity (in ampere) </li>
<li><b>Sub_metering_1</b>: energy sub-metering No. 1 (in watt-hour of active energy). It corresponds to the kitchen, containing mainly a dishwasher, an oven and a microwave (hot plates are not electric but gas powered). </li>
<li><b>Sub_metering_2</b>: energy sub-metering No. 2 (in watt-hour of active energy). It corresponds to the laundry room, containing a washing-machine, a tumble-drier, a refrigerator and a light. </li>
<li><b>Sub_metering_3</b>: energy sub-metering No. 3 (in watt-hour of active energy). It corresponds to an electric water-heater and an air-conditioner.</li>
</ol>

## Loading the data





When loading the dataset into R, please consider the following:

* The dataset has 2,075,259 rows and 9 columns. First
calculate a rough estimate of how much memory the dataset will require
in memory before reading into R. Make sure your computer has enough
memory (most modern computers should be fine).

```
About 140MB
```

* We will only be using data from the dates 2007-02-01 and
2007-02-02. One alternative is to read the data from just those dates
rather than reading in the entire dataset and subsetting to those
dates.

* You may find it useful to convert the Date and Time variables to
Date/Time classes in R using the `strptime()` and `as.Date()`
functions.

* Note that in this dataset missing values are coded as `?`.


This is how we read the data:

```
power.data <- read.table(
    './data/household_power_consumption.txt',
    header = T,        # the file has a header
    sep=";",           # the separator is ;
    na.strings = '?'   # the NA values are ?
)
```

We then add a new column in which we store a datetime value,
obtained from converting the `$Date` and `$Time` *string* columns
into *actual datetime* values:

```
power.data$DateTime <- strptime(paste(power.data$Date, power.data$Time), format="%d/%m/%Y %T")
```

-----


## Plots

The following is constructed using the four plot files from the 
course work. 

If you want to make this *R Markdown* file work, you will want 
to download the file mentioned above, unzip it,
and put it in a `data` subfolder before you `knittr` this *Rmd* file.

Or you could just take a look at the rendered [README.html](README.hhtml) file.

**NOTE:** if you run `knitter` (or `Ctrl-Shift-K` in RStudio) on this *Rmd* file,
expect to take a some amount of time as it loads that really large file in memory.

### Step 0

Set options to prevent the files from auto-executing (the course required it),
then `source` each file when it comes to plotting the results.

```{r}
# none of the plots should auto-execute
options(plot1.run=F, plot2.run=F, plot3.run=F, plot4.run=F)
# source the data-reading file
source('dataRunner.R')
# data is in:
dfile <- './data/household_power_consumption.txt'
# will output to:
outFile = NA

# only reading this once
data <- readData(dfile, date.subset = c('1/2/2007', '2/2/2007'))
```

All the figures you see below are generated live.
(If you read the README.html, then they are embedded within that file)

### Plot 1

```{r}
source('plot1.R')
plot1(data, outFile = outFile)
```



### Plot 2

```{r}
source('plot2.R')
plot2(data, outFile = outFile)
```


### Plot 3

```{r}
source('plot3.R')
plot3(data, outFile = outFile)
```


### Plot 4

```{r}
source('plot4.R')
plot4(data, outFile = outFile)
```

